---
title: "Fit 2-Way Complete Model"
description: "Fit a two-way complete (factorial) model and test which effects to keep."
categories: [Experiments, "Model Fitting"]
---

::: {.callout-note title="Introduction"}
Fit a two-way complete model with both main effects and their interaction, then
test which terms to retain.
:::

```{r}
#| message: false
#| warning: false

# read data
paperTowel <-
  read.table(
    "../../data/paper.towel.strength.txt",
    header = TRUE
  )

# consider as factors
paperTowel$A <- as.factor(paperTowel$A)
paperTowel$B <- as.factor(paperTowel$B)
paperTowel$C <- as.factor(paperTowel$C)

# fit model including AB interaction
model7.8 <- aov(strength ~ A + B + C + A*B, data=paperTowel)
```

$H_0$: An effect is insignificant to the model

$H_a$: An effect is not insignificant and should be added to the model

I will use the ANOVA table to determine P-values found via F testing.

Decision Rule: Any p-value under 0.01 will suggest that the corresponding
effect is significant and should be maintained in the model. I will use 0.01,
so that after the four tests, the end result will have a confidence of 0.04
as desired in the problem statement.

```{r}
# create anova table
drop1(model7.8, ~., test="F")
```

Only factors B and C should be included in the model, since both have
p-values < 0.01 via F testing. Factor A and the interaction between factors
A and B are considered insignificant to this model.


## See also

- [Impossible to Interpret](impossible-to-interpret.qmd)
- [Interaction Terms](interaction-terms.qmd)

*Data: `paper.towel.strength.txt` — see [sources & provenance](../../data/README.md#experiments).*
