---
title: "Impossible to Interpret (Limited Sample Size)"
description: "See how too few observations and empty cells leave factorial effects unmeasurable."
categories: [Experiments, "Experimental Design"]
---

::: {.callout-note title="Introduction"}
A cautionary example: with empty cells and too few observations, some factorial
effects simply cannot be estimated.
:::

```{r}
#| message: false
#| warning: false
pacman::p_load(dplyr, kableExtra)

# paperTowel carried over from the Fit 2-Way Complete Model task
paperTowel <- read.table("../../data/paper.towel.strength.txt", header = TRUE)
paperTowel$A <- as.factor(paperTowel$A)
paperTowel$B <- as.factor(paperTowel$B)
paperTowel$C <- as.factor(paperTowel$C)

# read data
towel2 <- filter(paperTowel, order <11)

# fit model
model7.13 <- aov(strength ~ A + B + C + A*B, data=towel2)
```

Is it possible to perform an analysis of variance of these data, using a
model that includes main effects and the AB interaction as required by the
experimenters? If so, analyze the experiment.

```{r}
# anova table
drop1(model7.13, ~., test="F")
```

With only ten observations, we are not able to conclude any effects belong in
the model, main effects or interaction.

Use a computer program to fit a three-way complete model. Can all of the main
effects and interactions be measured? If not, investigate which models could
have been used in the analysis of such a design with two empty cells and
unequal numbers of observations in the other cells.

```{r}
model7.13b <- aov(strength ~ A*B*C, data=towel2)

kable(drop1(model7.13b, ~., test="F")) %>%
        kable_styling(font_size = 7)
```

No, not all of the effects can be measured. We can consider models that
include Factor A, Factor B, both Factors A and B, or both Factor A and B as
well as their interaction effect.


## See also

- [Fit 2-Way Complete Model](fit-2-way-complete-model.qmd)

*Data: `paper.towel.strength.txt` — see [sources & provenance](../../data/README.md#experiments).*
