---
title: "Interaction Terms"
categories: [Experiments, "Interaction Terms"]
---

```{r}
#| message: false
#| warning: false
pacman::p_load(dplyr, ggplot2)

# read data
weld <- read.table("../../data/weld.strength.txt", header=TRUE)

weld$gage <- as.factor(weld$gage)                 # gage as factor
weld$time <- as.factor(weld$time)                 # time as factor

weld_model <- aov(strength ~ gage * time,         # two-way complete model
                  data=weld)
anova(weld_model)                                 # anova table

# determine cut-off for decision rule
qf(.95, 8, 15)                                   # (alpha, (a-1)(b-1), ab)
```

### Test for Interaction Term

$H_0^{AB}: (\alpha\beta)_{ij}-(\overline{\alpha\beta})_{i.}-(\overline{\alpha\beta})_{j.}+(\overline{\alpha\beta})_{..} = 0$

$H_a^{AB}: (\alpha\beta)_{ij}-(\overline{\alpha\beta})_{i.}-(\overline{\alpha\beta})_{j.}+(\overline{\alpha\beta})_{..} \neq 0$

Decision Rule:
I will reject the null hypothesis if $\frac{msAB}{msE} > F_{(a-1)(b-1,\, n-ab)}$
in which case the corresponding p-value will be less than 0.05.

Conclusion:
Because 6.826 > 2.640797 and the corresponding p-value is 0.0007, I reject
the null hypothesis and conclude that the interaction between gauge level and
weld time is not negligible.

```{r}
#| message: false
#| warning: false
attach(weld)
interaction.plot(gage, time, strength,            # interaction plot
                 type="b",
                 main="Interaction Plot Between Gage Setting and Time")
```

The plot supports my conclusion that interaction has an effect. I see that
levels 1 and 3 for time of weld cross the other plots of time, which is
generally indicative of interaction.

**Should we consider effect of gage bar setting?**

If we determine that interaction is not negligible, as is the case for this
study, we should proceed with a cell means model and consider possible
treatments rather than levels of an individual factor. This seems well-suited
for this study in particular, because we wish to determine the procedure which
results in the strongest weld.

Instead of considering main effects individually, we would consider the
"treatment" of gauge and time as one effect as shown below...

```{r}
#| message: false
#| warning: false

weld2 <- weld %>%
  mutate(Treatment = ifelse(trtmt == 11, 1,         # rename treatments
                            ifelse(trtmt == 12, 2,
                                   ifelse(trtmt == 13, 3,
                                          ifelse(trtmt == 14, 4,
                                                 ifelse(trtmt == 15, 5,
                                                        ifelse(trtmt == 21, 6,
                                                               ifelse(trtmt == 22, 7,
                                                                      ifelse(trtmt == 23, 8,
                                                                             ifelse(trtmt == 24, 9,
                                                                                    ifelse(trtmt == 25, 10,
                                                                                           ifelse(trtmt == 31, 11,
                                                                                                  ifelse(trtmt == 32, 12,
                                                                                                         ifelse(trtmt == 33, 13,
                                                                                                                ifelse(trtmt == 34, 14, 15)))))))))))))))
ggplot(weld2,
       aes(x = Treatment, y = strength, group=Treatment)) +
  geom_boxplot()                                    # view treatment boxplots
```

Predictably, the treatments with the highest weld strength belong to those
with the gage level at 1 or 2 and the time level of 3. These were also shown
to be highest on the interaction plot.
