---
title: "Linear and Quadratic Trends"
description: "Test for linear and quadratic trends across ordered treatment levels with polynomial contrasts."
categories: [Experiments, "Inference & Prediction"]
---

::: {.callout-note title="Introduction"}
Use polynomial trend contrasts to test whether the response changes linearly or
quadratically across ordered treatment levels.
:::

```{r}
#| message: false
#| warning: false
pacman::p_load(lsmeans)

# read data
water <- read.table("../../data/water.boiling.txt", header=TRUE)

water$salt <- as.factor(water$salt)                 # as factor
water$burner <- as.factor(water$burner)             # as factor

water_model <- aov(time ~ salt + burner,            # main effects model
                   data=water)
```

### Linear Trend

$H_0$: there is no linear trend

$H_a$: there is a linear trend

Decision Rule: If p < 0.01, I will reject the null hypothesis and conclude
there is a linear trend.

```{r}
#| message: false
#| warning: false

water_model_means <- lsmeans(water_model, ~salt)    # find least square means
linear_summary <- summary(contrast(water_model_means,

                                   # linear trend contrasts
                                   list(`Linear Trend` = c(-3, -1, 1, 3)),

                                   adjust = "tukey"), #tukey method
                          infer=c(T,T),
                          level=.99,               # 99% family-wise confidence
                          side="two-sided")
linear_summary
```

Because 0 is contained in the confidence interval and the p-value is greater
than 0.01, I reject that there exists a linear trend between teaspoons of salt
added to water and time to boil.

### Quadratic Trend

$H_0$: there is no quadratic trend

$H_a$: there is a quadratic trend

Decision Rule: If p < 0.01, I will reject the null hypothesis and conclude
there is a quadratic trend.

```{r}
#| message: false
#| warning: false

quad_summary <- summary(contrast(water_model_means,

                                 # quadratic trend contrasts
                                 list(`Quadratic Trend` = c(1, -1, -1, 1)),

                                 adjust = "tukey"), # tukey method
                        infer=c(T,T),
                        level=.99,                  # 99% family-wise confidence
                        side="two-sided")
quad_summary
```

Because 0 is contained in the confidence interval and the p-value is greater
than 0.01, I reject that there exists a quadratic trend between teaspoons of
salt added to water and time to boil.


## See also

- [Family-wise Confidence Intervals](familywise-confidence-intervals.qmd)
- [Choosing Contrasts](choosing-contrasts.qmd)

*Data: `water.boiling.txt` — see [sources & provenance](../../data/README.md#experiments).*
