---
title: "Check Assumptions"
description: "Check normality, constant variance, independence, and outliers for a linear model."
categories: [Regression, Diagnostics]
---

::: {.callout-note title="Introduction"}
Run these residual plots and tests to verify the core linear-model assumptions
before trusting any inference from the fit.
:::

Assumptions to check:

1. Normality
2. Constant Variance
3. Independence
4. Outliers

```{r}
#| message: false

# read data
gpa <- read.table("../../data/CH03PR03.txt", header=TRUE)

# plot data
plot(gpa$X,gpa$Y, ylab="Freshman GPA", xlab="ACT Score", main="Scatter Plot", pch=18, col="orange")

# plot normality
hist(gpa$X, xlab="ACT Scores", ylab="Frequency of Score", main="Histogram", col="orange")

# plot residuals
gfit <- lm(Y ~ X, data=gpa) #fit linear model
gpa.e <- resid(gfit) #residuals
plot(gpa$X, gpa.e, xlab="ACT Score", ylab="Residual", main="Residual Plot", col="orange", pch=18)
abline(h=0) #horizontal line
abline(h=1, lty=c(2)) #dotted line
abline(h=-1, lty=c(2)) #dotted line

# plot QQ plot
library(car)
qqPlot(rstandard(gfit), col="orange", pch=18)

# test normality (reject if < 0.05)
shapiro.test(gpa.e)
```

::: callout-note
## Extra
View residuals against other variables to see if they should be included in
the linear model.
:::

```{r}
# plot residuals against unused variables
plot(gpa$V2, gpa.e, col="orange", pch=18, xlab="V2", ylab="Residuals", main="V2 Residual Plot")
abline(h=0)
plot(gpa$V3, gpa.e, col="orange", pch=18, xlab="V3", ylab="Residuals", main="V3 Residual Plot")
abline(h=0)
```


## See also

- [Simple Linear Regression](simple-linear-regression.qmd)
- [Outliers & Influential Points](outliers-influential-points.qmd)
- [Check Assumptions](../experiments/check-assumptions.qmd)

*Data: `CH03PR03.txt` — see [sources & provenance](../../data/README.md#regression).*
