---
title: "Indicator Variables"
description: "Add a categorical predictor to a regression with an indicator (dummy) variable."
categories: [Regression, "Model Fitting"]
---

::: {.callout-note title="Introduction"}
Use an indicator variable to let a single regression capture the shift between
two categories (coded 0/1).
:::

```{r}
# read data: gpa1 (original) + gpa2 (indicator) = gpa (full dataset)
gpa1 <- read.table("../../data/CH01PR19-1.txt", header=TRUE)
gpa2 <- read.table("../../data/CH08PR16.txt", header=TRUE)
gpa <- cbind(gpa1, gpa2[1]) #combine

# fit linear model
gfit <- lm(GPA ~ ACT + Indicator, data=gpa)
summary(gfit)

# confidence interval for beta_2
qt <- qt(.975, 117)
ll <- -.0943 - (qt* .11997)
ul <- -.0943 + (qt* .11997)
```

The 95% confidence interval for $\beta_2$ is (`r ll`, `r ul`). We fail to
reject the null hypothesis, because 0 is contained in the confidence interval.
This suggests there is no value added by including $X_2$ in the model.

```{r}
# plot residuals against variable; color by indicator
gpa.e <- resid(gfit)
palette(c("gray", "red"))
test <- gpa$Indicator + 1
plot(gpa$ACT, gpa.e, col=test, pch=18,
     main="Residual Plot Against ACT Score Separated by Indicator",
     xlab="ACT Score", ylab="Residuals")
```


## See also

- [Interaction Terms](interaction-terms.qmd)

*Data: `CH01PR19-1.txt`, `CH08PR16.txt` — see [sources & provenance](../../data/README.md#regression).*
