---
title: "Interaction Terms"
description: "Let one predictor's effect depend on another using an interaction term."
categories: [Regression, "Model Fitting"]
---

::: {.callout-note title="Introduction"}
Add an interaction term when the slope for one predictor should change with the
level of another.
:::

```{r}
# data carried over from the Indicator Variables task
gpa1 <- read.table("../../data/CH01PR19-1.txt", header=TRUE)
gpa2 <- read.table("../../data/CH08PR16.txt", header=TRUE)
gpa <- cbind(gpa1, gpa2[1])

#fit linear model
gfit_int <- lm(GPA ~ ACT + Indicator + ACT*Indicator, data=gpa)
summary(gfit_int)
```

When no major field of concentration is indicated, the interaction term is
dropped. However, when a major field of concentration *is* indicated by the
student, the interaction coefficient contributes to the slope of the
regression line pertaining to students who had indicated their preference of
major, such that for one unit increase in ACT score, the student's freshman
GPA increased by 0.059488.

```{r}
# test interaction term not = 0
t_star1 <- .062245/.026487
t1 <- qt(.975, 116)
```

$H_0: \beta_3 = 0$

$H_a: \beta_3 \neq 0$

Because t\*, `r t_star1`, is greater than t, `r t1`, we can reject the null
hypothesis and conclude that the interaction term adds value to the
regression function.


## See also

- [Indicator Variables](indicator-variables.qmd)

*Data: `CH01PR19-1.txt`, `CH08PR16.txt` — see [sources & provenance](../../data/README.md#regression).*
