---
title: "Multiple Logistic Regression"
categories: [Regression, "Multiple Logistic Regression"]
---

A local health clinic sent fliers to its clients to encourage everyone, but
especially older persons at high risk of complications, to get a flu shot in
time for protection against an expected flu epidemic. In a pilot follow-up
study, 159 clients were randomly selected and asked whether they actually
received a flu shot. A client who received a flu shot was coded $Y=1$, and a
client who did not receive a flu shot was coded $Y=0$. In addition, data were
collected on their age ($X_1$), and their health awareness. The latter data
were combined into a health awareness index ($X_2$), for which higher values
indicate greater awareness. Also included in the dta was client gender, where
males were coded $X_3=1$ and females were coded $X_3=0$.

Multiple logistic regression model with three predictor variables in first-
order terms is assumed to be appropriate.

```{r}
#| message: false
#| warning: false

# read data
vaxx <- read.table("../../data/CH14PR14.txt", header=TRUE)

# fit logistic model
gfit3 <- glm(Y ~ X1 + X2 + X3, data=vaxx, family = binomial(link ="logit"))
summary(gfit3)
```

The fitted response function is
$$\hat{\pi} = \frac{\exp(-1.17716 + 0.07279X_1 - 0.09899X_2 + 0.43397X_3)}{1+\exp(-1.17716 + 0.07279X_1 - 0.09899X_2 + 0.43397X_3)}$$

```{r}
# find the odds ratio for each of the variables and interpret
exp(gfit3$coefficients[2])
exp(gfit3$coefficients[3])
exp(gfit3$coefficients[4])
```

It is expected that when all other variables are held constant, one unit
increase in age increases the probability of having gotten a flu shot by
7.55%.

It is expected that when all other variables are held constant, one unit
increase in a person's health awareness score decreased the probability of
them having gotten a flu shot by nearly 10%.

It is expected that when all other variables are held constant, a man is 54%
more likely to have gotten a flu shot than a woman with the same age and
health awareness.

```{r}
# find probability a 55 year old male with a health index of 60 receives vaccine
pi = (exp(-1.17716 + 0.07279*55 - 0.09899*60 + 0.43397*1))/(1+exp(-1.17716 + 0.07279*55 - 0.09899*60 + 0.43397*1))

# confidence interval: males with ages differing 30 and index differing 25
exp(30*gfit3$coefficients[2])
exp(25*gfit3$coefficients[3])
test1 <- exp(30*(confint.default(gfit3, level=.95)))
test2 <- exp(25*(confint.default(gfit3, level=.95)))
```

The 90% joint confidence intervals for the odds ratios are (`r test1[2,1]`,
`r test1[2,2]`) for male clients whose ages differ by 30 years and
(`r test2[3,1]`, `r test2[3,2]`) for male clients whose health awareness
index differs by 25.

Reviewing these intervals, I am 90% confident that the probability of
receiving a flu shot increases if a subject is 30 years older and decreases
if a subject has a health awareness index 25 points higher.

```{r}
# wald test for inclusion of x3: alpha = 0.05
norm <- qnorm(.975)                             # cutoff for decision
summary(gfit3)                                  # p-value
z_star <- .43397/.52179                         # z* statistic
```

$H_0: \beta_3 = 0$

$H_a: \beta_3 \neq 0$

Decision Rule: If the test statistic is greater than `r norm`, then I will
reject the null hypothesis and conclude that $X_3$ should be included in the
regression model.

Conclusion: The z\* statistic is `r z_star`, which is less than `r norm`;
therefore, I fail to reject the null hypothesis and do not have evidence to
support including $X_3$ in the model.

P-value: The p-value is 0.40558, which is greater than $\alpha$. This
supports my conclusion.

```{r}
#| message: false
#| warning: false

# likelihood ratio test for inclusion of x3: alpha = 0.05
chi4c <- qchisq(.95, df=1)                      # chi squared cutoff
library(lmtest)                                 # required package
nested <- glm(Y ~ X1 + X2,                      # model without x3
              data=vaxx,
              family = binomial(link ="logit"))
summary(nested)
lrtest(nested, gfit3)                           # test
```

NOTE: the Likelihood Ratio Tests are the same as Deviance Statistic-based
tests

Full Model: $[1+\exp(-X'\beta_F)]^{-1}$
where $X'\beta_F =  -1.17716 + 0.07279X_1 - 0.09899X_2 + 0.43397X_3$

Reduced Model: $[1+\exp(-X'\beta_R)]^{-1}$
where $X'\beta_R = 1.45778 + 0.07787X_1 - 0.09547X_2$

$H_0: \beta_3 = 0$

$H_a: \beta_3 \neq 0$

Decision Rule: If the test statistic is greater than `r chi4c`, then I will
reject the null hypothesis and conclude that $X_3$ should be included in the
regression model.

Conclusion: The test statistic, $G^2$, is 0.7022. Because it is less than
`r chi4c`, I fail to reject the null hypothesis and do not have enough
evidence to include $X_3$ in the model.

P-value: The p-value is 0.402, which supports my conclusion.

Comparison to Wald: Both tests conclude that $X_3$ does not improve the
regression model. Their p-values were very close, both supporting this
conclusion.

```{r}
#| message: false
#| warning: false

# likelihood ratio test for inclusion of second-order terms: alpha = 0.05
chi4d <- qchisq(.95, df=3)                      # chi squared cutoff

x_1sq <- vaxx$X1^2                              # x1 squared term
x_2sq <- vaxx$X2^2                              # x2 squared term
x_3sq <- vaxx$X3^2                              # x3 squared term
interact12 <- vaxx$X1*vaxx$X2                   # interaction of x1 and x2
interact13 <- vaxx$X1*vaxx$X3                   # interaction of x1 and x3
interact23 <- vaxx$X2*vaxx$X3                   # interaction of x2 and x3

complex <- glm(vaxx$Y ~ vaxx$X1 + vaxx$X2 + vaxx$X3 + x_1sq + x_2sq + interact12,
               family = binomial(link ="logit"))

summary(complex)
lrtest(gfit3, complex)                          # test
```

Full Model: $[1+\exp(-X'\beta_F)]^{-1}$
where $X'\beta_F = 19.6040080 - 0.0331617X_1 - 0.7529410X_2 + 0.6200280X_3 - 0.0005456X_1^2 + 0.0040101X_2^2 + 0.0033484X_1X_2$

Reduced Model: $[1+\exp(-X'\beta_R)]^{-1}$
where $X'\beta_R = -1.17716 + 0.07279X_1 - 0.09899X_2 + 0.43397X_3$

$H_0: \beta_4 = \beta_5 = \beta_6 = 0$

$H_a:$ At least one of $\beta_4, \beta_5,$ or $\beta_6 = 0$

Decision Rule: If the test statistic is greater than `r chi4d`, then I will
reject the null hypothesis and conclude that the second-order terms should be
included in the regression model.

Conclusion: The test statistic, $G^2$, is 2.1263. Because it is less than
`r chi4d`, I fail to reject the null hypothesis and do not have enough
evidence to support adding the second-order terms.

P-value: The p-value is 0.5466, which supports my conclusion.

```{r}
#| message: false
#| warning: false

# model selection with forward BIC
fit0 <- glm(Y ~ 1, family=binomial(), data=vaxx)  # null model

fitfull <- glm(vaxx$Y ~ vaxx$X1 + vaxx$X2 + vaxx$X3 +  interact12 + interact13 + interact23,
               family = binomial(link ="logit"))  # all first order terms and interactions

fitbic <- step(fit0,
               scope=list(upper=fitfull),
               data=vaxx,
               direction="forward",
               k=log(nrow(vaxx)))

# deviance residuals vs. estimated probabilities
fitbic <- glm(vaxx$Y ~ vaxx$X1 + vaxx$X2 + interact12,
              family = binomial(link ="logit"))  # model selected using forward BIC
r1 <- loess(residuals(fitbic) ~ predict(fitbic)) # fit residuals against fitted pi's
y <- predict(r1, se=TRUE)                        # predict residual

xvals <- predict(fitbic)                         # linear predictors of fitbic
orderx <- order(predict(fitbic))                 # order of lowest to highest xvals
summary(fitbic)                                  # print summary


plot(predict(fitbic), residuals(fitbic),         # predicted pi vs. residual
     col="slateblue1", pch=16,
     ylab="Deviance Residuals",
     xlab="Fitted Probabilities",
     main="Residuals Plot")
abline(h=0, lty=2, col="grey", lwd=2)            # grey line on y = 0
lines(x=xvals[orderx],                           # linear predictors, ordered
      y=y$fit[orderx],                           # corresponding predicted residual
      lwd=2)                                     # loess line
```

I have some concern about the adequacy of the model fit due to the departure
of the loess line from an expected value of zero for fitted probabilities
above zero.
