---
title: "Simple Logistic Regression"
description: "Model a binary outcome from one predictor and interpret the odds ratio."
categories: [Regression, "Model Fitting"]
---

::: {.callout-note title="Introduction"}
Use simple logistic regression to model a binary outcome from a single
predictor, reading the slope as an odds ratio.
:::

```{r}
#| message: false
#| warning: false

# plot logistic model with b0 = 20, b1 = -.2
X <- cbind(1:200)
pi_1a <- exp(20-.2*X)/(1+exp(20-.2*X))
plot(X, pi_1a, pch=18, col="slateblue1", main="Logistic Mean Response with Given Parameters")

# odds ratio is approximately exp(b1)
odds126 <- exp(20-.2*126)/(1+exp(20-.2*126))
odds125 <- exp(20-.2*125)/(1+exp(20-.2*125))
ratio <- odds126/odds125
exp(-.2)

# read data
dues <- read.table("../../data/CH14PR07.txt", header=TRUE)

# fit logistic model
gfit2 <- glm(Renewal ~ Increase, data=dues, family = binomial(link ="logit"))
summary(gfit2)
```

The fitted response function is
$$\hat{\pi} = \frac{\exp(-4.80751 + 0.12508X)}{1+\exp(-4.80751 + 0.12508X)}$$

```{r}
#| message: false
#| warning: false

# generate prediction curve
predIncrease <- seq(0,100, by=1)                    # create x values
logit.hat2 <- predict.glm(gfit2,                    # ???
                          newdata = data.frame(Increase=predIncrease),
                          se.fit=TRUE)
prob.hat2 <- exp(logit.hat2$fit)/(1+exp(logit.hat2$fit)) # logistic equation

# plot original data
plot(dues$Increase,
     dues$Renewal,
     pch=18, col="slateblue1",
     main="Annual Dues Increase",
     ylab="Renewal Selection",
     xlab="Increase Amount")

# plot loess smoothed line
loess2 <- loess.smooth(dues$Increase, dues$Renewal)
lines(loess2, col="slateblue2")                     # loess smooth line
lines(predIncrease, prob.hat2, col="red")           # predicted values line

# calculate and interpret odds ratio
beta_1 <- exp(gfit2$coefficients[2])
```

The value is 1.133237 and suggests that for each additional dollar dues
increase, the odds of a member choosing not to renew their dues increases 13%.

```{r}
dues_b <- gfit2$coefficients                        # save b1 and b0

# estimated probability that members will not renew if the dues increase $40
pi_2d <- exp(dues_b[1]+dues_b[2]*40)/(1+exp(dues_b[1]+dues_b[2]*40))

# dues increase for which 75 percent of the members are expected not to renew
Y <- exp(dues_b[1]+dues_b[2]*X)/(1+exp(dues_b[1]+dues_b[2]*X)) # equation
answer <- X[which(Y > .74 & Y < 0.76)]              # find Y = .75
```


## See also

- [Multiple Logistic Regression](multiple-logistic-regression.qmd)

*Data: `CH14PR07.txt` — see [sources & provenance](../../data/README.md#regression).*
